Constructing Statistical Deformable Models with Shape Flow

### 2015-04-22 contents/evaluation.tex
Convergence Test
Updates from ShareLaTeX
Introduction Updated by Stefanos
Temp Push to share with tassos

To evaluate the quality of model,

\subsection{Using Sparse Point Annotation}

\subsection{Using Drawing Annotation}

>>>>

In this section we evaluate the performance of the proposed methodology in a variety of tasks. Results for 4 different experiments are reported. Experiment \ref{exp:1} compares the fitting accuracy of Dense Active Appearance Models (DAAMs) built using our methodology with respect to that of classic AAMs on the problems of non-rigid face and ear alignment in-the-wild. Experiment \ref{exp:2} investigates the impact that the Support Vector Shape representation has on the performance achieved by DAAMs on the the first experiment. Experiment \ref{exp:3} evaluates the effectiveness of the proposed methodology to build DAAMs of objects that are difficult to annotated with respect to a consistent set of landmarks, such as bottles and human bodies. Finally, experiment \ref{exp:4} serves as a qualitative demonstration of how the proposed pipeline can use be effectively used to generate novel modified instances of an object, e.g. caricatures, from simple hand drawn sketches.

\subsection{Shape representation and flow constraints}

\begin{figure}[t!]
\centering
\includegraphics[width=0.38\textwidth]{resources/Fig_Alignment/svs_experiment}
\caption{CED over 68 landmarks on the LFPW database.}
\label{fig:face_ced}
\end{figure}

\begin{table}[t]
\small
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\emph{Method}         & \emph{mean $\pm$ std} & \emph{median} & $\leq 0.05$\\
\hline
Initialization        & $\pm$ & & \%\\
\hline
SVS representation    & $\pm$ & & \%\\
Binary representation & $\pm$ & & \%\\
\hline
Constraint Flow       & $\pm$ & & \%\\
Unconstrained Flow    & $\pm$ & & \%\\
\hline
\end{tabular}
\caption{Fitting statistics on the LFPW database.}
\label{tab:face_statistics}
\end{table}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{resources/Fig_SVS/faces_svs_binary_gaussian}
        \caption{AAM Principle Components Contribution}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_SVS/ears_svs_binary_gaussian}
        % \caption{Shape Flow Principle Components Contribution}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_SVS/faces_svs_binary_gaussian_statistics}
        % \caption{AAM Principle Components Contribution}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.2\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_SVS/ears_svs_binary_gaussian_statistics}
        % \caption{Shape Flow Principle Components Contribution}
    \end{subfigure}
    \caption{Principle Components Contribution}
\end{figure}

\subsection{Non-rigid object alignment in-the-wild}
\label{exp:1}

\subsubsection{Face alignment}

This experiment compares the fitting accuracy of our DAAMs and classic AAMs. For this experiment, both models were built using the 811 training images of the Labelled Faces Parts in-the-Wild (LFPW) \cite{} database and corresponding 68 point annotations provided by the iBUG group \footnote{\label{ibug_300}\url{http://ibug.doc.ic.ac.uk/resources/300-W/}}. Classic AAMs were built using the standard procedure described in \cite{} using piece-wise affine as their motion model. Our DAAMs were built using the methodology described on the main part of the paper. Notice that we modified the previous 68 point annotations of all images by adding and removing $\pm$5 to and from the that were not or subtracted $\approx$

Results for this experiment are reported over the 224 test images of the LFPW cite{} database. 66
points ground truth landmark annotations were provided by
the iBUG group8
. All methods were initialized by perturbing
the ground truth scale and translation parameters with
Gaussian noise (rotations were not considered) and applying
the resulting transformation to mean of the shape model.
(Notice that this procedure produces initializations that are
considerably more challenging than those reported in the recent
AAM literature [32, 33]). The Cumulative Error Distributions
(CED) for this experiment is shown in Figures 1
and 2. Figures 3 and 4 shows the evolution of the mean
normalized point-to-point error as a function of the number
of iterations run by each algorithm

\begin{figure}[t!]
\centering
\includegraphics[width=0.38\textwidth]{resources/Fig_Alignment/face_experiment}
\caption{CED over 68 landmarks on the LFPW database.}
\label{fig:face_ced}
\end{figure}

\begin{table}[t]
\small
\centering
\begin{tabular}{|l|c|c|c|}
\hline
\emph{Method}   & \emph{mean $\pm$ std} & \emph{median} & $\leq 0.05$\\
\hline
Initialization  & $\pm$ & & \%\\
\hline
INT - dAAM      & $\pm$ & & \%\\
INT - AAM       & $\pm$ & & \%\\
HOG - dAAM      & $\pm$ & & \%\\
HOG - AAM       & $\pm$ & & \%\\
\hline
\end{tabular}
\caption{Fitting statistics on the LFPW database.}
\label{tab:face_statistics}
\end{table}

%\begin{figure}[h!]
%    \centering
%    \begin{subfigure}[b]{0.2\textwidth}
%            \includegraphics[width=\textwidth]{resources/Fig_Alignment/face_intensity}
%        % \caption{AAM Principle Components Contribution}
%    \end{subfigure}
%    \hfill
%    \begin{subfigure}[b]{0.2\textwidth}
%            \includegraphics[width=\textwidth]{resources/Fig_Alignment/ear_intensity}
%        % \caption{Shape Flow Principle Components Contribution}
%    \end{subfigure}
%    \\
%    \begin{subfigure}[b]{0.2\textwidth}
%            \includegraphics[width=\textwidth]{resources/Fig_Alignment/face_intensity_statistics}
%        % \caption{AAM Principle Components Contribution}
%    \end{subfigure}
%    \hfill
%    \begin{subfigure}[b]{0.2\textwidth}
%            \includegraphics[width=\textwidth]{resources/Fig_Alignment/ear_intensity_statistics}
%        % \caption{Shape Flow Principle Components Contribution}
%    \end{subfigure}
%    \caption{CED Plot and Statistics of Comparison with Classic AAM}
%\end{figure}

\subsubsection{Ear alignment}

\subsubsection{Objects with loosely define landmarks}

\subsection{DAAMs without consistent annotations}
\label{exp:3}

\subsection{Object generation from sketches}
\label{exp:4}

\begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_01}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.12\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_01_base}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_01_aam}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_01_of}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_02}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.12\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_02_base}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_02_aam}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_02_of}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_03}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.12\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_03_base}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_03_aam}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_03_of}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_04}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.12\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_04_base}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_04_aam}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.1\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Draw/test_04_of}
    \end{subfigure}
    \caption{Figure}
    \label{fig:draw}
\end{figure}

### 2015-04-22 contents/implementation.tex
Edited Intro
Figure Added for Introduction and Implementation
Temp Push to share with tassos

@@

\begin{figure}[h!]
    \centering
        \includegraphics[width=0.5\textwidth]{resources/architecture}
    \caption{Training Architecture}
    \label{fig:archi}
\end{figure}

@@

We propose a novel framework for building deformable models that does not require any consistent set of annotated landmarks and is based on dense shape representation. Our method only requires a set of point or curve annotation that does not need to be consistent over different training samples. It combines NICP, SVS representation and multi-image subspace flow in an effective framework that has significant descriptive power.

The proposed pipeline is depicted in Figure .... It takes as input a set of training images of an object class, along with point and/or curve annotations. Its first step is the application of ICP to achieve an initial alignment of the training images. Next, the shapes that correspond to the training data are consistently represented using multichannel SVS. These are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all shapes of the training set. In order for the flow estimation to be accurate and exploit the correlation across the different training shapes, it utilises a correspondence basis that is built on sparse correspondences given by Nonrigid ICP alignment of the annotated data. Finally, the dense correspondences that are yielded by the optical flow serve as automatic dense landmarks and are used by a dense AAM training and fitting procedure. The following sections discuss in more detail the steps of the proposed pipeline.

>>>>

\begin{figure*}[t!]
    \centering
        \includegraphics[width=\textwidth]{resources/architecture}
    \caption{Training Architecture}
    \label{fig:archi}
\end{figure*}

@@

@@

We propose a novel framework for building deformable models that does not require any consistent set of annotated landmarks and is based on dense shape representation. Our method only requires a set of point or curve annotation that does not need to be consistent over different training samples. It combines Non-Rigid Iterative closest point (NICP) \cite{}, Support Vector Shape (SVS) \cite{} representation and multi-image subspace flow \cite{} in an effective framework that has significant descriptive power.

The proposed pipeline is depicted in Figure~\ref{fig:archi}. It takes as input a set of training images of a particular object class, along with (possibly inconsistent) point and/or curve annotations. Its first step is the application of ICP to achieve an initial alignment of the training images. Next, the shapes that correspond to the training data are consistently represented using multichannel SVS. These are then fed into a multi-image subspace flow estimation that establishes dense correspondences between all shapes of the training set. In order for the flow estimation to be accurate and exploit the correlation across the different training shapes, it utilises a correspondence basis that is built on sparse correspondences given by Nonrigid ICP alignment of the annotated data. Finally, the dense correspondences that are yielded by the optical flow serve as automatic dense landmarks and are used by a dense AAM training and fitting procedure. The following sections discuss in more detail the steps of the proposed pipeline.

### 2015-04-22 contents/introduction.tex
Temp Push to share with tassos

% \clearpage

Building and fitting statistical deformable models of objects is a well-studied and popular area in the intersection of computer vision and machine learning [..]. Recently, we have witnessed tremendous developments in building and fitting models of human face and body using images captured in unconstrained recording conditions, usually referred to as "in-the-wild" [..]. This is attributed to 
\begin{itemize}

\item the abundance of complex visual data, spread mostly through the internet via web services such as Youtube, Flickr and Google Images. The latter has led to the development of huge databases of "in-the-wild" images of human faces and bodies [..]. Equally important is the task of manual annotation of images with regards to face and body parts and that has been undertaken by several research teams [..].

\item the development of powerful visual features, able to describe objects and its parts in a robust manner (e.g., Scale Invariant Feature Transforms (SIFT) [39], Histogram of Oriented Gradients (HoGs) [16] and recently Deep Convolutional Neural Networks [..]), as well as generative and discriminative methodologies for learning deformable models.

@@

\item annotating with regards to consistent landmark annotation is extremely time consuming, laborious and labour intensive work [..], which usually can be performed by a trained person. Furthermore, for many objects it requires a highly skilled person to identify and annotate a set of images with regards to a consistent set of landmarks. For example human ears have very complicated inner structures (e.g. helix, crus antihelicis, scapha, tragus, lobe etc.) which differs remarkably between ears. Furthermore, 
certain ear parts, such as  Fossa Triangularies and Crus Helicis, do not appear in all ears and  their visibility is highly sensitive to pose and illumination variation. Please see Fig...

\item the nature of many deformable objects do not allow to be consistently annotated with regards to a set of landmarks (e.g. bottles, fruits etc.). Also the boundary/outline of objects, such as faces, ears, body can be very difficultly annotated consistently. That is why many state-of-the-art opt to leave the boundary out for comparison [..]. An example is given in Fig.. The majority of the state-of-the-art methods for landmark-based deformable models [..] are not applicable for objects with no consistent set of landmarks.

>>>>

% \clearpage
\begin{figure}[t!]
    \centering
    \begin{subfigure}[b]{0.145\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_0_0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.145\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_0_1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.145\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_0_2}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_1_0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_1_1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_1_2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_1_3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_1_4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_1_5}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.07\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_2_0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.07\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_2_1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.07\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_2_2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.07\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_2_3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.07\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_2_4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.07\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_2_5}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_3_0}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_3_1}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_3_2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_3_3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_3_4}
    \end{subfigure}
   	\hfill
    \begin{subfigure}[b]{0.075\textwidth}
            \includegraphics[width=\textwidth]{resources/Fig_Intro/intro_3_5}
    \end{subfigure}
    \caption{Demonstration of classical landmark annotation and curve annotation of}
    \label{fig:intro}
\end{figure}

Building and fitting statistical deformable models of objects is a well-studied and popular area in the intersection of computer vision and machine learning [..]. Recently, we have witnessed tremendous developments in building and fitting models of human face and body using images captured in unconstrained recording conditions, usually referred to as "in-the-wild" [..]. This is attributed to: 
\begin{itemize}

\item The abundance of complex visual data, spread mostly through the internet via web services such as Youtube, Flickr and Google Images. The latter has led to the development of large databases of "in-the-wild" images of human faces and bodies [..]. Equally important is the task of manual annotation of images with regards to face and body parts and that has been undertaken by several research teams [..].

\item The development of powerful visual features, able to describe objects and its parts in a robust manner (e.g., Scale Invariant Feature Transforms (SIFT) [39], Histogram of Oriented Gradients (HoGs) [16] and recently Deep Convolutional Neural Networks [..]), as well as generative and discriminative methodologies for learning deformable models.

@@

\item Annotating with regards to consistent landmark annotation is extremely time consuming, tedious and labour intensive work [..], which usually can be performed by a trained person. Furthermore, for many objects it requires a highly skilled person to identify and annotate a set of images with regards to a consistent set of landmarks. For example human ears have very complicated inner structures (e.g. helix, crus antihelicis, scapha, tragus, lobe etc.) which differs remarkably between ears. Furthermore, certain ear parts, such as  Fossa Triangularies and Crus Helicis, do not appear in all ears and  their visibility is highly sensitive to pose and illumination variation.

\item The nature of many deformable objects does not allow them to be consistently annotated with regards to a set of landmarks (e.g. bottles, fruits etc.). Also the boundary/outline of objects, such as faces, ears, body can be very difficultly annotated consistently, Fig. \ref{fig:main}. That is why many state-of-the-art opt to leave the boundary out for comparison [..]. An example is given in Fig.. The majority of the state-of-the-art methods for landmark-based deformable models [..] are not applicable for objects with no consistent set of landmarks.
